prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis

### 2022-10-06 inst/MyArticle/paper.Rmd
updating comparison table.

```{r tableOverview}
xx <- list(
  MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "Y"),
  ROPECA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y"),
  limma  = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y"),
  MSqRob2_rlm = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y", md = "?"),
  proDA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y", md = "Y"),
  prolfqua = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y",eb = "Y", md = "Y"))

bb <- data.frame(dplyr::bind_rows(xx))
rownames(bb) <- names(xx)
knitr::kable(data.frame(bb), caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling. Y - yes, * - repeated measurements are modeled using a fixed effect. ? - the hurdle model was published but is not available in the msqrob2 package.")
```

@@

knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, TN - true negatives, P - all positive cases (all $E.~coli$ proteins), N - all negative cases (all $H.~sapiens$ proteins), m - all proteins.")
```

@@

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
## load peptide level data
d <- prolfqua::prolfqua_data('data_ionstar')$filtered()

## create R6 obejct
lfqd <- prolfqua::LFQData$new(d$data, d$config)

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
# specify differences among groups
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
)
## fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
models <- prolfqua::build_model(lfqp, modelFunction)

@@

# merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::addContrastResults(prefer = contr, add = conI)

@@

getpath <- function(filN){
    f1 <- paste0("../../inst/Benchresults/",filN)
    if(f1 == ""){
        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
    }
    message(f1)
    return(f1)
}
allBenchmarks <- readRDS(getpath("allBenchmarks.RDS"))
benchmark_msstats <- readRDS(getpath("benchmark_msstats.RDS"))
benchmark_prodA <- readRDS(getpath("benchmark_medpolish_proDA.RDS"))
msFragger <- readRDS(getpath("MSFragger_medpol_benchmark.RDS"))

allBenchmarks$benchmark_mssstats <- benchmark_msstats
allBenchmarks$benchmark_msFragger <- msFragger
allBenchmarks$benchmark_proDA <- benchmark_prodA
allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA"   )]

@@

```{r benchmarkROC}
ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})

res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})

res <- res |> dplyr::mutate(whatfix = dplyr::case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res |> dplyr::group_by(contrast,whatfix) |>
    dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
res <- dplyr::inner_join(res, norm)
res <- dplyr::mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

@@

nrgg <- dds |> ggplot2::ggplot(ggplot2::aes(x = name, y = nrcontrasts )) + 
  ggplot2::geom_bar(stat = "identity", fill="white", colour = "black") + 
  ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 200, max(dds$nrcontrasts) + 10)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
  ggplot2::geom_text(ggplot2::aes(label = paste0(round(nrcontrasts, digits = 1), paste0("  (",round(percent, digits = 1),"%)"))),
            vjust = 0, hjust = -0.2, angle = -90) #+ 
pB <- nrgg + ggplot2::labs(tag = "B")

@@

colnames(xM) <- c("prolfqua functions", "model")
knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions that can be used to fit various models.",
             format = "latex", booktabs=TRUE) |>
    kableExtra::kable_styling(latex_options="scale_down") |> 
    kableExtra::add_footnote(label = "in development")

@@

knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))
```

@@

# column with protein abundances
atable$setWorkIntensity("total.intensity")

>>>>

```{r tableOverview}
caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling. Y - yes, * - repeated measurements are modeled using a fixed effect."

xx <- list(
  MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "Y"),
  ROPECA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y"),
  limma  = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y",eb = "Y"),
  msqrob2 = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y", eb = "Y", md = "Y"),
  proDA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y", md = "Y"),
  prolfqua = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y",eb = "Y", md = "Y"))

bb <- data.frame(dplyr::bind_rows(xx))
rownames(bb) <- names(xx)
knitr::kable(data.frame(bb), format = "latex", caption = caption)
```

@@

knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, TN - true negatives, P - all positive cases (all $E.~coli$ proteins), N - all negative cases (all $H.~sapiens$ proteins), m - all proteins.", format = "latex" )
```

@@

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
## load peptide level data
d <- prolfqua::old2new(prolfqua::prolfqua_data('data_ionstar')$filtered())
d <- prolfqua::LFQData$new(d$data, d$config)
## create R6 obejct
lfqd <- prolfqua::LFQData$new(d$data, d$config)

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
# specify differences among groups
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
)
## fit model
lmmodel <- paste(lfqp$response()," ~ dilution.")
modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
models <- prolfqua::build_model(lfqp, modelFunction)

@@

# merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)

@@

getpath <- function(filN){
    f1 <- paste0("../../inst/Benchresults/",filN)
    if (f1 == "") {
        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
    }
    message(f1)
    return(f1)
}
allBenchmarks <- readRDS(getpath("allBenchmarks.RDS"))
benchmark_msstats <- readRDS(getpath("benchmark_msstats.RDS"))
benchmark_prodA <- readRDS(getpath("benchmark_medpolish_proDA.RDS"))
msFragger <- readRDS(getpath("MSFragger_medpol_benchmark.RDS"))
benchmark_msqrob <- readRDS(getpath("benchmark_msqrob.RDS"))

allBenchmarks$benchmark_mssstats <- benchmark_msstats
allBenchmarks$benchmark_msFragger <- msFragger
allBenchmarks$benchmark_proDA <- benchmark_prodA
allBenchmarks$benchmark_msqrob <- benchmark_msqrob

allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA", "benchmark_msqrob")]

@@

```{r benchmarkROC}
ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
res <- res |> dplyr::mutate(whatfix = dplyr::case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res |> dplyr::group_by(contrast, whatfix) |>
    dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
res <- dplyr::inner_join(res, norm)
res <- dplyr::mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

@@

nrgg <- dds |> ggplot2::ggplot(ggplot2::aes(x = name, y = nrcontrasts )) + 
  ggplot2::geom_bar(stat = "identity", fill = "white", colour = "black") + 
  ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 200, max(dds$nrcontrasts) + 10)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
  ggplot2::geom_text(ggplot2::aes(label = paste0(round(nrcontrasts, digits = 1), paste0("  (",round(percent, digits = 1),"%)"))),
            vjust = 0, hjust = -0.2, angle = -90) #+ 
pB <- nrgg + ggplot2::labs(tag = "B")

@@

colnames(xM) <- c("prolfqua functions", "model")
knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions that can be used to fit various models.",
             format = "latex", booktabs = TRUE) |>
    kableExtra::kable_styling(latex_options = "scale_down") |> 
    kableExtra::add_footnote(label = "in development")

@@

knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))

@@

# column with protein abundances
atable$set_work_intensity("total.intensity")
